\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\title{Advent of Code: Day 6 Solution}
\date{}
\begin{document}
\maketitle
\noindent Each fish reproduces at the end of their own internal timer, and
resets its own timer to 6.\\

\noindent Define \(T(L,D)\) be to the number of fish after \(D\) days
starting with a single fish having an internal timer of \(L\). We have three
cases
\begin{enumerate}
    \item [Case 1.] \(D = 0, L \in \mathbb{N}\). There are no more days
          remaining, this fish cannot reproduce. We still have only 1 fish.
    \item [Case 2.] \(L = 0, D > 0\). \(D\) is at minimum 1, so this fish
          can reproduce. This gives us another fish with a timer set to 8 and the
          original fish's timer is reset to 6. Thus in \(D - 1\) days we will have
          \(T(8, D-1) + T(6, D - 1)\)
    \item [Case 3.] The fish is not yet ready to reproduce, we skip until
          either the fish timer expires again or the days have been exhausted -
          whichever comes first. This is \(T(L - \min(L,D), D - \min(L,D))\)
\end{enumerate}
We get the following recurrence,
\begin{eqnarray}
    T(L,D) = \begin{cases}
        1                               & \textrm{if } D = 0, L \in \mathbb{N} \\
        T(8, D - 1) + T(6, D - 1)       & \textrm{if } L = 0, D > 0            \\
        T(L - \min(L,D), D - \min(L,D)) & \textrm{otherwise}
    \end{cases}
\end{eqnarray}
\begin{enumerate}
    \item [Base Case.] \(k = 0, 0 \leq L \leq 8\).\\
          Then \(T(L,k) = T(L,0) = 1\) as wanted.
    \item[I.H] Suppose by induction that
        \begin{equation}
            P(n): T(L,n) \textrm{ gives the number of fish after } n \textrm{ days for any } 0 \leq L \leq 8
        \end{equation}
        holds whenever \(0 \leq n < k\).
    \item [W.T.S] \(P(k)\) holds. \(T(L,k)\) has 3 cases,
          \begin{enumerate}
              \item \(k = 0\). Then \(T(L,k) = T(L,0) = 1\) as wanted (by base
                    case).
              \item \(L = 0, k > 0\). \(T(L,k) = T(8, k -1) + T(6, k - 1)\).\\
                    Since \(k - 1 < k\), by I.H, \(T(8, k - 1)\) returns the correct
                    value.\\
                    Since \(k - 1 < k\), by I.H, \(T(6, k - 1)\) returns the correct
                    value.\\
                    It is easy to see that number of fish after \(k\) days will the be
                    the sum of these two values. Therefore \(T(L,k) = T(8, k -1) + T(6,
                    k - 1)\) as wanted.
              \item \(L > 0, k > 0\). Then we have 3 cases,
                    \begin{enumerate}
                        \item[\(L < k\):] Then it follows that
                            \begin{align}
                                T(L,k) & = T(L - \min(L,k), k - \min(L,k))    \\
                                       & =T(0,k') \textrm{ where } k' = k - L
                            \end{align}
                            Since \(k' = k - L < k\), \(T(0,k')\) returns the correct value.
                            (By I.H)
                        \item[\(L > k\):] Then it follows that
                            \begin{align}
                                T(L,k) & = T(L - \min(L,k), k - \min(L,k))    \\
                                       & =T(L',0) \textrm{ where } L' = L - k
                            \end{align}
                            Since \(0 < k\), \(T(L',0)\) returns the correct value. (By I.H)
                        \item[\(L = k\):] Then it follows that
                            \begin{align}
                                T(L,k) & = T(L - \min(L,k), k - \min(L,k)) \\
                                       & =T(0,0)                           \\
                                       & = 1
                            \end{align}
                            Since \(0 < k\), \(T(0,0)\) returns the correct value. (By I.H)

                    \end{enumerate}
          \end{enumerate}
\end{enumerate}
\end{document}